### 2022-06-16 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@ec18cd6940b2ac05776c605d198846cc9cd2f92c 🚀

Based on the previous result, we extracted the yellow component of the
risk map (TP add methods), to provide a single measure of risk varying
between 0 and 1. This measure is presented in fig.~\ref{fig:risk}.
However, this maps the potential risk, which must be weighed by the
potential for contacts with humans. As a proxy for this measure, we used
the proportion of build/urban land from the EarthEnv dataset: this is a
reasonable proxy for the density of humans per unit area, which
increases the probability of pathogen spread more widely (Hazarie et
al., 2021). Since human activity is required to amplify the frequency of
virus encounters and thus create areas of viral amplification, mapping
the potential risk against measures of land use is required to generate
a more actionable assessment of risk. This map is presented in
fig.~\ref{fig:compound}. Most of South America and Europe are at low
risk, as although densely populated, settlements tend to be in areas
with lower potential risk. However, this mapping reveals that South-East
Asia, the Indian subcontinent, and parts of sub-Saharan Africa, are at
high risk due to the overlap between built areas and bat communities
representing more opportunities for cross-species transmission of
betacoronaviruses.

@@

Our study focuses largely on the biogeography of hosts. Yet, we know
that viruses with high host plasticity, that is, the ability of a given
virus to adapt to various taxonomic orders and ecological groups
(Kreuder Johnson et al., 2015); are more likely to amplify viral
spillover, followed by secondary human-to-human transmission, and
geographical spread (Hazarie et al., 2021). High viral host plasticity
is an especially important trait for RNA viruses such as betacov
(Kreuder Johnson et al., 2015; Haddad et al., 2021). Indeed, our
analysis of viral sequences reveals that Latin America is a hotspot of
viral distinctiveness, suggesting that this part of the bats-betacov
system may be undergoing independent evolutionary dynamics (related
species sharing viruses that are different from the rest of the global
pool). The other hotspot of viral distinctiveness is S.E. Asia, in which
richness is high but sharing is low; this suggests a different type of
evolutionary dynamics (unrelated viruses coevolving with evolutionarily
distinct hosts, generating high diversity locally). Both of these areas
should be priority areas for sampling, especially since
(\textbf{Becker?}) advance that they harbor undiscovered hosts of
beta-coronaviruses. This diversity of hosts, and the mechanisms by which
the exchange of viruses occurs between species, is largely affected by
the local environmental conditions and environmental change.

There are several factors that drive changes in the diversity of bats
(Alves et al., 2018), but human activities' effects on the ecosystem
(like modifications of land use) could significantly decrease it.
Therefore, it can be suggested that changes in the diversity of betacovs
in bats are linked to their biogeographic variation, and human
population density and other anthropogenic factors are decisive
moderators for its implications in public health. With the increase of
contact between humans and potential hosts, we also increase the risk of
emergence of novel diseases (Johnson et al., 2020), as previous studies
on RNA viruses suggest the importance of host phylogeography at the time
of virus dispersal (Gryseels et al., 2017). One of these scenarios where
interaction between bats and humans can occur can be seed dispersal in
tropical agroecosystems. It opens the discussion of whether the fruits
thrown by bats not only disperse seeds but could also be a source of
indirect interaction between viruses of bat origin and humans (Deshpande
et al., XX). This represents a challenge for conservation strategies and
disease ecology since we have areas with both potential for the
acquisition of zoonotic viruses and bat-human interaction; in
particular, the challenge lies in the fact that actual exposure must be
quantified from several scenarios, including both direct and indirect
bat - human interaction.

@@

@@

@@

@@

@@

>>>>

Based on the previous result, we extracted the yellow component of the
risk map (TP add methods), to provide a single measure of risk varying
between 0 and 1. This measure is presented in fig.~\ref{fig:risk}.
However, this maps the potential risk, which must be weighed by the
potential for contacts with humans. As a proxy for this measure, we used
the proportion of build/urban land from the EarthEnv dataset: this is a
reasonable proxy for the density of humans per unit area, which
increases the probability of pathogen spread more widely (Hazarie et al.
2021). Since human activity is required to amplify the frequency of
virus encounters and thus create areas of viral amplification, mapping
the potential risk against measures of land use is required to generate
a more actionable assessment of risk. This map is presented in
fig.~\ref{fig:compound}. Most of South America and Europe are at low
risk, as although densely populated, settlements tend to be in areas
with lower potential risk. However, this mapping reveals that South-East
Asia, the Indian subcontinent, and parts of sub-Saharan Africa, are at
high risk due to the overlap between built areas and bat communities
representing more opportunities for cross-species transmission of
betacoronaviruses.

@@

Our study focuses largely on the biogeography of hosts. Yet, we know
that viruses with high host plasticity, that is, the ability of a given
virus to adapt to various taxonomic orders and ecological groups
(Kreuder Johnson et al. 2015), are more likely to amplify viral
spillover, followed by secondary human-to-human transmission, and
geographical spread (Hazarie et al. 2021). High viral host plasticity is
an especially important trait for RNA viruses like betacoronaviruses
(Haddad et al. 2021). Indeed, our analysis of viral sequences reveals
that Latin America is a hotspot of viral distinctiveness, suggesting
that this part of the bats-betacoronavirus system may be undergoing
independent evolutionary dynamics (related species sharing viruses that
are different from the rest of the global pool). The other hotspot of
viral distinctiveness is S.E. Asia, in which richness is high but
sharing is low; this suggests a different type of evolutionary dynamics
(unrelated viruses coevolving with evolutionarily distinct hosts,
generating high diversity locally). Both of these areas should be
priority areas for sampling, especially since Becker et al. (2022)
advance that they harbor undiscovered hosts of beta-coronaviruses. This
diversity of hosts, and the mechanisms by which the exchange of viruses
occurs between species, is largely affected by the local environmental
conditions and environmental change.

There are several factors that drive changes in the diversity of bats
(Alves et al. 2018), but human activities' effects on the ecosystem
(like modifications of land use) could significantly decrease it.
Therefore, it can be suggested that changes in the diversity of betacovs
in bats are linked to their biogeographic variation, and human
population density and other anthropogenic factors are decisive
moderators for its implications in public health. With the increase of
contact between humans and potential hosts, we also increase the risk of
emergence of novel diseases (Johnson et al. 2020), as previous studies
on RNA viruses suggest the importance of host phylogeography at the time
of virus dispersal (Gryseels et al. 2017). One of these scenarios where
interaction between bats and humans can occur can be seed dispersal in
tropical agroecosystems. It opens the discussion of whether the fruits
thrown by bats not only disperse seeds but could also be a source of
indirect interaction between viruses of bat origin and humans (Deshpande
et al. 2022). This represents a challenge for conservation strategies
and disease ecology since some areas can haveboth potential for the
acquisition of zoonotic viruses and bat-human interactions; in
particular, the challenge lies in the fact that actual exposure must
then be quantified accounting for several transmission scenarios,
including both direct and indirect bat - human interaction.

@@

\leavevmode\hypertarget{ref-Alves2018GeoVar}{}%
Alves, D. M. C. C. et al. 2018. Geographic variation in the relationship
between large-scale environmental determinants and bat species richness.
- Basic and Applied Ecology 27: 1--8.

@@

\leavevmode\hypertarget{ref-Deshpande2022ForFru}{}%
Deshpande, K. et al. 2022. Forbidden fruits? Ecosystem services from
seed dispersal by fruit bats in the context of latent zoonotic risk. -
Oikos (Copenhagen, Denmark): oik.08359.

@@

\leavevmode\hypertarget{ref-Gryseels2017WheVir}{}%
Gryseels, S. et al. 2017. When Viruses Don't Go Viral: The Importance of
Host Phylogeographic Structure in the Spatial Spread of Arenaviruses (JH
Kuhn, Ed.). - PLOS Pathogens 13: e1006073.

\leavevmode\hypertarget{ref-Haddad2021SarPos}{}%
Haddad, D. et al. 2021. SARS-CoV-2: Possible recombination and emergence
of potentially more virulent strains (H Attoui, Ed.). - PLOS ONE 16:
e0251368.

\leavevmode\hypertarget{ref-Hazarie2021IntPop}{}%
Hazarie, S. et al. 2021. Interplay between population density and
mobility in determining the spread of epidemics in cities. -
Communications Physics 4: 191.

@@

\leavevmode\hypertarget{ref-Johnson2020GloShi}{}%
Johnson, C. K. et al. 2020. Global shifts in mammalian population trends
reveal key predictors of virus spillover risk. - Proceedings of the
Royal Society B: Biological Sciences 287: 20192736.

@@

\leavevmode\hypertarget{ref-KreuderJohnson2015SpiPan}{}%
Kreuder Johnson, C. et al. 2015. Spillover and pandemic properties of
zoonotic viruses with high host plasticity. - Scientific Reports 5:
14830.
